Designing Types for R, Empirically

### 2020-05-16 paper/main.tex
stuff
minor

% Contracts can be removed by calling \code{remove\_contract}.  Contracts can
% be selectively enabled or disabled for code blocks by wrapping them in calls
% to \code{ignore\_contracts} and \code{capture\_contracts}.  These functions
% return a data frame that contains all the information about failed and
% successful contract assertions in the wrapped code blocks.  Multiple
% failures can introduce noise; this is avoided by setting
% \code{contractr.severity} to \code{'silence'}.  \contractr still performs
% type-checking so that results can be obtained by calling
% \code{get\_contracts}. Setting \code{contractr.severity} flag to
% \code{'error'} turns the type-checking warnings to errors and
% halt the program.

\paragraph{Checking the values} A number of properties can be checked
by a simple tag check, namely whether a value is a: null, environment, vector
or scalar. Other properties require an inspection of the contents of a
value: the absence of \NA or the type of the elements of list. Union types
require checking for each of the alternatives. We do not check higher-order
functions in this version of the tool.
When argument values are wrapped in promises (this is not always the case
due to compiler optimizations), in order to retain the non-strict semantics
of R, the expression held in the promise is wrapped in a call to the type
checker, and type checking is delayed until the promise is forced. This
leads to corner cases such that the type checking of a function may happen
after that function has returned.
Return values require care as well.  Functions return the last expression
they evaluate, thus to avoid having to analyze the code of the called
function the checker will register a callback on the exit hook. The hook is
executed in the function call's environment. Another wrinkle is due to
longjumps which causes active function calls on the stack to be
discarded. When they are discarded, their exit hooks are called.  But, these
function do not have a return value to type-check. \contractr deals with
this problem by allocating a unique sentinel object which serves as the
return value for calls that are discarded. The exit hook does not call the
type-checker if it see the sentinel.

@@

\begin{figure}[!h]\centering\includegraphics[width=.8\linewidth]
  {plots/corpus.pdf}\caption{Corpus}\label{fig:corpus}
\end{figure}

>>>>

\paragraph{Checking the values} A number of properties can be checked
by a simple tag check, namely whether a value is a: null, environment, vector
or scalar. Other properties require an inspection of the contents of a
value: the absence of \NA or the type of the elements of list. Union types
require checking for each of the alternatives. We do not check higher-order functions in this version of the tool. Checking
them could be done by dynamically wrapping functions with checks. The usage
of higher-order functions remains limited in R. When argument values are wrapped in promises (this is not always the case
due to compiler optimizations), in order to retain the non-strict semantics
of R, the expression held in the promise is wrapped in a call to the type
checker, and type checking is delayed until the promise is forced. This
leads to corner cases such that the type checking of a function may happen
after that function has returned.

Return values require care as well.  Functions return the last expression
they evaluate, thus to avoid having to analyze the code of the called
function the checker will register a callback on the exit hook. The hook is
executed in the function call's environment. Another wrinkle is due to
longjumps which causes active function calls on the stack to be
discarded. When they are discarded, their exit hooks are called.  But, these
function do not have a return value to type-check. \contractr deals with
this problem by allocating a unique sentinel object which serves as the
return value for calls that are discarded. The exit hook does not call the
type-checker if it see the sentinel.

@@

\begin{figure}[!h]\centering\includegraphics[width=.9\linewidth]
  {plots/corpus.pdf}\caption{Corpus}\label{fig:corpus}
\end{figure}
